String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-05-11 2023-1-19/sections/abstract.tex
intro
space adjustment

Regular expressions (abbreviated as regex) and string-length function are widely used in string-manipulating programs. 
%Solving string constraints involving regular membership and string-length function has been widely investigated. 
Counting is a frequently used feature of regular expressions that count the number of matchings of sub-patterns. The state-of-the-art string solvers e.g. CVC4/5, Z3str3(RE), and Z3seq are incapable of solving string constraints with regex-counting and string-length efficiently, especially when the counting bounds are large. In this work, we propose an automata-theoretic approach for solving such a class of string constraints efficiently. 
%
The main idea is to model the counting operators symbolically by registers in automata, instead of unfolding them explicitly, thus alleviating the state explosion problem resulted from the unfolding.  
%
Moreover, we model the string lengths by registers as well. 
%Since Parikh images of automata can be defined by linear integer arithmetic formulas, 
As a result, the satisfiability of string constraints with regex-counting and string-length is reduced to the satisfiability of linear integer arithmetic, which can then be solved by off-the-shelf SMT solvers. 
%
%we extract linear integer arithmetic formulas from automata, which encode the constraints on the values of registers, and utilize the off-the-shelf SMT solvers to solve the linear integer arithmetic constraints. 
%
To improve the performance further, on the one hand, we propose techniques to reduce automata sizes, 
%thus reducing the sizes of linear arithmetic formulas computed from automata as well, 
and on the other hand, we utilize under approximations to search for solutions.   
% we propose state reduction techniques and under-approximation to improve the performance further. 
We implement the algorithms and do extensive experiments to validate our approach. The experimental results show that 
our approach can solve more problem instances than the other tools, with the same or a better performance.
% performance of this approach compared against the state-of-the-art string solvers.

@@

\keywords{Regex Expressions, Counting, Length, Automata, Registers, Linear Integer Arithmetic}
%\end{document}

>>>>

Regular expressions (abbreviated as regex) and string-length function are widely used in string-manipulating programs. 
%Solving string constraints involving regular membership and string-length function has been widely investigated. 
Counting is a frequently used feature of regular expressions that count the number of matchings of sub-patterns. The state-of-the-art string solvers e.g. CVC4/5, Z3str3(RE), and Z3seq are incapable of solving string constraints with regex-counting and string-length efficiently, especially when the counting bounds are large. In this work, we propose an automata-theoretic approach for solving such a class of string constraints efficiently. 
%
The main idea is to model the counting operators symbolically by registers in automata, instead of unfolding them explicitly, thus alleviating the state explosion problem resulted from the unfolding.  
%
Moreover, we model the string lengths by registers as well. 
%Since Parikh images of automata can be defined by linear integer arithmetic formulas, 
As a result, the satisfiability of string constraints with regex-counting and string-length is reduced to the satisfiability of linear integer arithmetic, which can then be solved by off-the-shelf SMT solvers. 
%
%we extract linear integer arithmetic formulas from automata, which encode the constraints on the values of registers, and utilize the off-the-shelf SMT solvers to solve the linear integer arithmetic constraints. 
%
To improve the performance further, on the one hand, we propose techniques to reduce automata sizes, 
%thus reducing the sizes of linear arithmetic formulas computed from automata as well, 
and on the other hand, we utilize under approximations to search for solutions.   
% we propose state reduction techniques and under-approximation to improve the performance further. 
We implement the algorithms and validate our approach on 48,834 benchmark instances. The experimental results show that 
our approach can solve more instances than the other tools, with comparable or better performance.
% performance of this approach compared against the state-of-the-art string solvers.

@@

%\keywords{Rgular Expressions, Counting, Length, Automata, Registers, Linear Integer Arithmetic}
%\end{document}

### 2023-05-11 2023-1-19/sections/implementation.tex
space adjustment

\paragraph*{AutomatArk benchmark suite.}
The AutomatArk benchmark suite is adapted from the AutomatArk suite in \cite{z3str3re} by filtering out the string constraints that contain no occurrences of counting operators and adding the length constraint $|x| > 10$ (where $x$ is a string variable). There are 8,215 instances in the AutomatArk suite here.
Note that the original AutomatArk benchmark suite in \cite{z3str3re} includes 19,979 instances in total, which are conjunctions of regular membership queries and generated out of regular expressions from \cite{automatark}. 
%Only 8,751 ones that contain counting operators out of 19,979 instances.

\paragraph*{RegCoL benchmark suite.} There are 40,628 RECL constraints in the RegCoL suite. These constraints are generated by extracting regular expressions with counting operators from \cite{regex_lingua_franca,redos_lenka} and constructing for each regular expression $e$ a RECL constraint of the form 
$$x \in e \wedge x \in \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*} \wedge |x| > 10,$$   
where $\Sigma^*(<+ >+'+''+\&)\Sigma^*$ is a regular expression that specifies at least one occurrence of special characters $<$, $>$, $'$, $''$, or $\&$, and $\overline{e'}$ denotes the complement of a regular expression $e'$. We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them.

@@

\begin{table}[ht]
\begin{center}
  \import{tables}{table_all.tex}
\end{center}
  \caption{Experiment results for Q1, where the timeout period is 60 seconds}
  \label{tab:results_all}
\end{table}

@@

\begin{table}[ht]
\begin{center}
  \subimport{tables}{table_heuristic.tex}
\end{center}
  \caption{Experiment results for Q2, where the timeout period is 60 seconds}
  \label{tab:results_heuristics}
\end{table}

>>>>

\medskip
\noindent
\emph{AutomatArk benchmark suite.}
The AutomatArk benchmark suite is adapted from the AutomatArk suite in \cite{z3str3re} by filtering out the string constraints that contain no occurrences of counting operators and adding the length constraint $|x| > 10$ (where $x$ is a string variable). There are 8,215 instances in the AutomatArk suite here.
Note that the original AutomatArk benchmark suite in \cite{z3str3re} includes 19,979 instances in total, which are conjunctions of regular membership queries and generated out of regular expressions from \cite{automatark}. 
%Only 8,751 ones that contain counting operators out of 19,979 instances.

\medskip
\noindent
\emph{RegCoL benchmark suite.} There are 40,628 RECL constraints in the RegCoL suite. These constraints are generated by extracting regular expressions with counting operators from \cite{regex_lingua_franca,redos_lenka} and constructing for each regular expression $e$ a RECL constraint of the form 
$x \in e \wedge x \in \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*} \wedge |x| > 10,$   
where $\Sigma^*(<+ >+'+''+\&)\Sigma^*$ is a regular expression that specifies at least one occurrence of special characters $<$, $>$, $'$, $''$, or $\&$, and $\overline{e'}$ denotes the complement of a regular expression $e'$. We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them.

@@

\begin{table}[t]
\vspace{-1mm}
\begin{center}
  \import{tables}{table_all.tex}
\end{center}
  \caption{Experiment results for Q1, where the timeout period is 60 seconds}
  \label{tab:results_all}
\vspace{-5mm}
\end{table}

\begin{table}[ht]
\vspace{-3mm}
\begin{center}
  \subimport{tables}{table_heuristic.tex}
\end{center}
  \caption{Experiment results for Q2, where the timeout period is 60 seconds}
  \label{tab:results_heuristics}
\vspace{-5mm}
\end{table}

@@

### 2023-05-11 2023-1-19/sections/introduction.tex
space adjustment

\paragraph*{Contribution} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur in regular expressions (called RECL for brevity). We propose an automata-theoretical approach for solving RECL constraints. Our main idea is to represent the counting operators by registers in automata, instead of unfolding them explicitly. Our approach is built upon an automata model called cost-enriched finite automata (CEFA), which was proposed in~\cite{atva2020} to solve string constraints involving the integer data type. Moreover, the string-length function can also be modeled by registers in CEFA. Therefore, the satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA.

According to the results from~ \cite{atva2020}, a linear integer arithmetic (LIA) formula can be computed to represent the potential values of registers in CEFA. Thus, the nonemptiness of CEFA can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers for LIA. 
%
Because the sizes of LIA formulas depend much on the sizes of CEFA, it is beneficial to reduce the sizes of CEFA, in order to achieve a good performance. As a result, we propose various techniques to reduce the state space, the number of transitions, and the number of registers for CEFA. Furthermore, since the computation of LIA formulas from CEFA is relatively expensive, we also propose a method of searching for strings that are accepted by CEFA, based on under approximations, before constructing the LIA formulas.

@@

At last, we would like to clarify the relationship of this work with~\cite{atva2020}, where CEFA was introduced. This work differs from \cite{atva2020} in the following aspects: In \cite{atva2020}, the regex-counting was unfolded explicitly. Moreover, in~\cite{atva2020}, the nonemptiness of CEFA w.r.t. LIA formulas was solved by utilizing the model checker nuXmv~\cite{nuxmv}. In contrast, this work solves it by integrating the techniques of Parikh image computation, automata-size reduction, and under approximation. The experimental results show that our approach solves 7,092 more instances while spending less time than the solver in~\cite{atva2020}. (See Table~\ref{tab:results_all}, where OSTRICH denotes the solver in \cite{atva2020}.)

\paragraph*{Related work} 
Here, we discuss more related work about regular expressions with counting and real-world regular expressions in programming languages.  
The determinism of regular expressions with counting was investigated in~\cite{GGM12,CL15}. Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to match a subclass of regular expressions with counting quickly. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regular expressions with counting.   Nevertheless, the nonemptiness problem of these automata was not considered in ~\cite{redos_lenka,HS+23,GKM23}, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length.
Real-world regular expressions in programming languages include features beyond classical regular expressions, e.g. the greedy/lazy Kleene star, capturing groups, and back references. Real-world regular expressions have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving \cite{CF+22}.

@@

\paragraph*{Organization} The rest of this paper is structured as follows: Section~\ref{sec:overview} gives an overview of the approach in this paper. Section~\ref{sec:pre} introduces the preliminaries. 
Section~\ref{sec:recl} presents the syntax and semantics of RECL. 
Section~\ref{sec:automaton} defines CEFA. Section~\ref{sec:algorithm} introduces the algorithm to solve RECL constraints. Section~\ref{sec:implementation} shows the experiment results. Section~\ref{sec:conclu} concludes this paper.

>>>>

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur in regular expressions (called RECL for brevity). We propose an automata-theoretical approach for solving RECL constraints. Our main idea is to represent the counting operators by registers in automata, instead of unfolding them explicitly. Our approach is built upon an automata model called cost-enriched finite automata (CEFA), which was proposed in~\cite{atva2020} to solve string constraints involving the integer data type. Moreover, the string-length function can also be modeled by registers in CEFA. Then the satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA with respect to a linear integer arithmetic (LIA) formula whose free variables are registers of CEFA.

According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA. Thus, the nonemptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers for LIA. 
%
Because the sizes of LIA formulas depend much on the sizes of CEFA, it is beneficial to reduce the sizes of CEFA, in order to achieve a good performance. As a result, we propose various techniques to reduce the state space and the number of transitions for CEFA. Furthermore, since the computation of LIA formulas from CEFA is relatively expensive, we also propose a method of searching for strings that are accepted by CEFA, based on under approximations, before constructing the LIA formulas.

@@

At last, we would like to clarify the relationship of this work with~\cite{atva2020}, where CEFA was introduced. This work differs from \cite{atva2020} in the following aspects: In \cite{atva2020}, regex-counting was unfolded explicitly. Moreover, in~\cite{atva2020}, the nonemptiness of CEFA w.r.t. LIA formulas was solved by utilizing the model checker nuXmv~\cite{nuxmv}. In contrast, this work solves it by integrating the techniques of Parikh-image computation, automata-size reduction, and under approximation. The experimental results show that our approach solves 7,092 more instances while spending less time than the solver in~\cite{atva2020}. (See Table~\ref{tab:results_all}, where OSTRICH denotes the solver in \cite{atva2020}.)

\medskip
\noindent
\emph{Related work.} 
Here, we discuss more related work about regular expressions with counting and real-world regular expressions in programming languages.  
The determinism of regular expressions with counting was investigated in~\cite{GGM12,CL15}. Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to match a subclass of regular expressions with counting quickly. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regular expressions with counting.   Nevertheless, the nonemptiness problem of these automata was not considered in ~\cite{redos_lenka,HS+23,GKM23}, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length.
Real-world regular expressions in programming languages include features beyond classical regular expressions, e.g. the greedy/lazy Kleene star, capturing groups, and back references. Real-world regular expressions have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving \cite{CF+22}.

@@

\medskip
\noindent
\emph{Organization.} 
The rest of this paper is structured as follows: Section~\ref{sec:overview} gives an overview of the approach in this paper. Section~\ref{sec:pre} introduces the preliminaries. 
Section~\ref{sec:recl} presents the syntax and semantics of RECL. 
Section~\ref{sec:automaton} defines CEFA. Section~\ref{sec:algorithm} introduces the algorithm to solve RECL constraints. Section~\ref{sec:implementation} shows the experiment results. Section~\ref{sec:conclu} concludes this paper.

### 2023-05-11 2023-1-19/sections/preliminaries.tex
space adjustment

\paragraph*{Strings and languages.}
A string over $\Sigma$ is a (possibly empty) sequence of elements from $\Sigma$,
denoted by $u, v, w, \ldots$. An empty string is denoted by $\varepsilon$. We use $\Sigma_\varepsilon$ to denote $\Sigma \cup \{\varepsilon\}$. We write $\Sigma^*$ (resp., $\Sigma^+$) for the set of all (resp. nonempty) strings over $\Sigma$.
%Let 
For a string $u$, we use $|u|$ to denote the number of letters in $u$. In particular, $|\varepsilon|=0$. 
Moreover, for $a \in \Sigma$, let $|u|_a$ denote the number of occurrences of $a$ in $u$. 
Assume $u=a_0\cdots a_{n-1}$ is nonempty and $i<j \in [0,n-1]$. %Then a \emph{position} of $u$ is a number $i \in [|u|]$ (Note that the first position is $1$, instead of  0). In addition, 
We let $u[i]$ denote $a_i$ and $u[i,j]$ for the substring %of $u$ starting from $i$ and ending with $j$ (i.e., 
$a_i\cdots a_j$. 
%\tl{check later for consistency}\zhilin{the indices start from 0, to be consistent with the semantics of substring and indexof}\zhilin{i do use it in section 4}
%
Let $u, v$ be two strings. We use $u \cdot v$ to denote the \emph{concatenation} of $u$ and $v$. A language $L$ over $\Sigma$ is a subset of strings.  
Let $L_1, L_2$ be two languages. Then the concatenation of $L_1$ and $L_2$, denoted by $L_1 \cdot L_2$,  is defined as $\{u \cdot v \mid u \in L_1, v \in L_2\}$. The union (resp. intersection) of $L_1$ and $L_2$, denoted by $L_1 \cup L_2$  (resp. $L_1 \cap L_2$), is defined as $\{u \mid u \in L_1 \mbox{ or } u \in L_2\}$ (resp. $\{u \mid u \in L_1 \mbox{ and } u \in L_2\}$). The complement of $L_1$, denoted by $\overline{L_1}$, is defined as $\{u \mid u \in \Sigma^*, u \not \in L_1\}$. 
The difference of $L_1$ and $L_2$, denoted by $L_1 \setminus L_2$, is defined as $L_1 \cap \overline{L_2}$. 
For a language $L$ and $n \in \Nat$, we define $L^n$ inductively as follows: $L^0= \{\varepsilon\}$, $L^{n+1} = L \cdot L^n$ for every $n \in \Nat$. 
Finally, define $L^* = \bigcup \limits_{n \in \Nat} L^n$.
%, that is, the string $w$ such that $|w|= |u| + |v|$ and for each $i \in [0, |u|-1]$, $w[i]= u[i]$, and for each $i \in [0,|v|-1]$, $w[|u|+i]=v[i]$.

\paragraph*{Finite automata.} 
A \emph{(nondeterministic) finite automaton} (NFA)  is a tuple $\NFA=(Q, \Sigma, \delta, I, F)$, where $Q$ is a finite set of states, $\Sigma$ is a finite alphabet, $\delta \subseteq Q \times \Sigma \times Q$ is the transition relation, $I \subseteq Q$ is the set of initial states, and $F \subseteq Q$ is the set of final states. For readability, we write a transition $(q, a, q') \in \delta$ as $q \xrightarrow[\delta]{a} q'$ (or simply $q \xrightarrow{a} q'$). %Moreover, when $\delta$ is clear from context, we omit $\delta$ in $q \xrightarrow[\delta]{a} q'$ and write $q \xrightarrow{a} q'$. 
The \emph{size} of an NFA $\NFA$, denoted by $|\NFA|$, is defined as the number of transitions of $\NFA$.
%
A \emph{run} of $\NFA$ on a string $w = a_1 \cdots a_n$ is a sequence of transitions $q_0 \xrightarrow{a_1} q_1 \cdots q_{n-1} \xrightarrow{a_n} q_n$ such that $q_0 \in I$. The run is \emph{accepting} if $q_n \in F$.
A string $w$ is accepted by an NFA $\NFA$ if there is an accepting run of $\NFA$ on $w$. In particular, the empty string $\varepsilon$ is accepted by $\NFA$ if $I \cap F \neq \emptyset$. The language of $\NFA$, denoted by $\Lang(\NFA)$, is the set of strings accepted by $\NFA$. 
%
An NFA $\NFA$ is said to be \emph{deterministic} if $I$ is a singleton and, for every $q \in Q$ and $a \in \Sigma$, there is at most one state $q' \in Q$ such that $(q, a, q') \in \delta$. We use DFA to denote deterministic finite automata. 
%

@@

\paragraph*{Linear integer arithmetic and Parikh images.}
We will also use standard quantifier-free/existential \emph{linear integer arithmetic} (LIA) formulae, which are typically ranged over by $\phi, \varphi$, etc. 
For a set $\mathfrak{X}$ of variables, we use $\Phi(\mathfrak{X})$ to denote the set of quantifier-free LIA formulas whose free variables are from $\mathfrak{X}$. 
For convenience, we use $\varphi(\vec{\anivar})$ with $\myvec{\anivar} = (\anivar_1, \cdots, \anivar_k)$ to denote an LIA formula $\varphi$ whose free variables are from $\{\anivar_1, \cdots, \anivar_k\}$. For an LIA formula $\varphi(\vec{\anivar})$ with $\myvec{\anivar} = (\anivar_1, \cdots, \anivar_k)$ and $\myvec{t} = (t_1, \cdots, t_k)$ be a tuple of integer terms, we use $\varphi[\myvec{t}/\myvec{\anivar}]$ to denote the formula obtained by replacing (simultaneously) $\anivar_i$ with $t_i$ for every $i \in [k]$.
%For convenience, we usually assume that a set $X$ of variables is a vector. 
%, which is essentially Presburger

>>>>

\medskip
\noindent \emph{Strings and languages.}
%\paragraph*{Strings and languages.}
A string over $\Sigma$ is a (possibly empty) sequence of elements from $\Sigma$,
denoted by $u, v, w, \ldots$. An empty string is denoted by $\varepsilon$. We use $\Sigma_\varepsilon$ to denote $\Sigma \cup \{\varepsilon\}$. We write $\Sigma^*$ (resp., $\Sigma^+$) for the set of all (resp. nonempty) strings over $\Sigma$.
%Let 
For a string $u$, we use $|u|$ to denote the number of letters in $u$. In particular, $|\varepsilon|=0$. 
Moreover, for $a \in \Sigma$, let $|u|_a$ denote the number of occurrences of $a$ in $u$. 
Assume $u=a_0\cdots a_{n-1}$ is nonempty and $i<j \in [0,n-1]$. %Then a \emph{position} of $u$ is a number $i \in [|u|]$ (Note that the first position is $1$, instead of  0). In addition, 
We let $u[i]$ denote $a_i$ and $u[i,j]$ for the substring %of $u$ starting from $i$ and ending with $j$ (i.e., 
$a_i\cdots a_j$. 
%\tl{check later for consistency}\zhilin{the indices start from 0, to be consistent with the semantics of substring and indexof}\zhilin{i do use it in section 4}
%
Let $u, v$ be two strings. We use $u \cdot v$ to denote the \emph{concatenation} of $u$ and $v$. A language $L$ over $\Sigma$ is a subset of strings.  
Let $L_1, L_2$ be two languages. Then the concatenation of $L_1$ and $L_2$, denoted by $L_1 \cdot L_2$,  is defined as $\{u \cdot v \mid u \in L_1, v \in L_2\}$. The union (resp. intersection) of $L_1$ and $L_2$, denoted by $L_1 \cup L_2$  (resp. $L_1 \cap L_2$), is defined as $\{u \mid u \in L_1 \mbox{ or } u \in L_2\}$ (resp. $\{u \mid u \in L_1 \mbox{ and } u \in L_2\}$). The complement of $L_1$, denoted by $\overline{L_1}$, is defined as $\{u \mid u \in \Sigma^*, u \not \in L_1\}$. 
The difference of $L_1$ and $L_2$, denoted by $L_1 \setminus L_2$, is defined as $L_1 \cap \overline{L_2}$. 
For a language $L$ and $n \in \Nat$, we define $L^n$ inductively as follows: $L^0= \{\varepsilon\}$, $L^{n+1} = L \cdot L^n$ for every $n \in \Nat$. 
Finally, define $L^* = \bigcup \limits_{n \in \Nat} L^n$.
%, that is, the string $w$ such that $|w|= |u| + |v|$ and for each $i \in [0, |u|-1]$, $w[i]= u[i]$, and for each $i \in [0,|v|-1]$, $w[|u|+i]=v[i]$.

\medskip
\noindent 
\emph{Finite automata.} 
A \emph{(nondeterministic) finite automaton} (NFA)  is a tuple $\NFA=(Q, \Sigma, \delta, I, F)$, where $Q$ is a finite set of states, $\Sigma$ is a finite alphabet, $\delta \subseteq Q \times \Sigma \times Q$ is the transition relation, $I \subseteq Q$ is the set of initial states, and $F \subseteq Q$ is the set of final states. For readability, we write a transition $(q, a, q') \in \delta$ as $q \xrightarrow[\delta]{a} q'$ (or simply $q \xrightarrow{a} q'$). %Moreover, when $\delta$ is clear from context, we omit $\delta$ in $q \xrightarrow[\delta]{a} q'$ and write $q \xrightarrow{a} q'$. 
The \emph{size} of an NFA $\NFA$, denoted by $|\NFA|$, is defined as the number of transitions of $\NFA$.
%
A \emph{run} of $\NFA$ on a string $w = a_1 \cdots a_n$ is a sequence of transitions $q_0 \xrightarrow{a_1} q_1 \cdots q_{n-1} \xrightarrow{a_n} q_n$ such that $q_0 \in I$. The run is \emph{accepting} if $q_n \in F$.
A string $w$ is accepted by an NFA $\NFA$ if there is an accepting run of $\NFA$ on $w$. In particular, the empty string $\varepsilon$ is accepted by $\NFA$ if $I \cap F \neq \emptyset$. The language of $\NFA$, denoted by $\Lang(\NFA)$, is the set of strings accepted by $\NFA$. 
%
An NFA $\NFA$ is said to be \emph{deterministic} if $I$ is a singleton and, for every $q \in Q$ and $a \in \Sigma$, there is at most one state $q' \in Q$ such that $(q, a, q') \in \delta$. We use DFA to denote deterministic finite automata. 
%

@@

\medskip
\noindent 
\emph{Linear integer arithmetic and Parikh images.}
We will also use standard quantifier-free/existential \emph{linear integer arithmetic} (LIA) formulae, which are typically ranged over by $\phi, \varphi$, etc. 
For a set $\mathfrak{X}$ of variables, we use $\Phi(\mathfrak{X})$ to denote the set of quantifier-free LIA formulas whose free variables are from $\mathfrak{X}$. 
For convenience, we use $\varphi(\vec{\anivar})$ with $\myvec{\anivar} = (\anivar_1, \cdots, \anivar_k)$ to denote an LIA formula $\varphi$ whose free variables are from $\{\anivar_1, \cdots, \anivar_k\}$. For an LIA formula $\varphi(\vec{\anivar})$ with $\myvec{\anivar} = (\anivar_1, \cdots, \anivar_k)$ and $\myvec{t} = (t_1, \cdots, t_k)$ be a tuple of integer terms, we use $\varphi[\myvec{t}/\myvec{\anivar}]$ to denote the formula obtained by replacing (simultaneously) $\anivar_i$ with $t_i$ for every $i \in [k]$.
%For convenience, we usually assume that a set $X$ of variables is a vector. 
%, which is essentially Presburger
